---
title: "LQ-PCA"
author: "Ryan Thomas"
date: "`r format(Sys.Date())`"
output: 
  md_document:
    variant: markdown_github
---
# About the data
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)

```

```{r packages, include=FALSE }
library(plyr)
library(rgdal)
library(sp)
library(dplyr)
library(data.table)
```



```{r getData, include=FALSE}
counties <- readOGR("https://datahub.io/dataset/e82a4e76-ccf0-4d3a-b2c8-eda2b52ab32a/resource/f2693c94-fd69-4baf-887b-f302a5188ead/download/counties.geojson",
                    "OGRGeoJSON")
counties@data$FIPS <- as.character(counties@data$FIPS)
for (i in 1:nrow(counties@data) ) {
  if ( nchar( counties@data$FIPS[i] ) == 4) {
    counties@data$FIPS[i] <- paste("0",as.character(counties@data$FIPS[i]), sep = "")
  } else {
     counties@data$FIPS[i] <- counties@data$FIPS[i]
  }
}
counties@data$FIPS <- as.factor(counties@data$FIPS)
str(counties@data$FIPS)
counties@data <- counties@data[,1:10]
```

```{r getEmploymentData, warning=FALSE}
sectors <- data.frame(cbind(
  c("11", "21","22", "23","31_33", "42","48_49", "51", "52","53", "54", "55", "56",
    "61", "62","71","72","81","92"),
  c("AgriForest",  "Mining", "Utilities","Construction", "Manufacturing",  "WholeTrade", 
    "RetailTrade", "Transportation",  "Information", "Finance", "RealEstate", "ProfScienceTech", 
    "Management",  "AdminWaste","Education","HealthCare","AccomodationFood","Other","PubAdmin"))
  )
names(sectors) <- c("sector", "description")

qcewGetIndustryData <- function (year, qtr, industry) {
  url <- "http://www.bls.gov/cew/data/api/YEAR/QTR/industry/INDUSTRY.csv"
  url <- sub("YEAR", year, url, ignore.case=FALSE)
  url <- sub("QTR", qtr, url, ignore.case=FALSE)
  url <- sub("INDUSTRY", industry, url, ignore.case=FALSE)
  read.csv(url, header = TRUE, sep = ",", quote="\"", dec=".", na.strings=" ", skip=0)
}
### Use this loop to go through the sectors loaded in the data frame 
for (i in 1:length(sectors[,1])) {
  if (i == 1) {
    print(paste("getting data for ", sectors[i,2]))
    dt1 <- qcewGetIndustryData("2015", "3", sectors[1,1])
    dt1 <- aggregate(dt1$month2_emplvl, by=list(Category=dt1$area_fips), FUN=sum)
    names(dt1) <- c("area_fips", as.character(sectors[1,2]))
    
  } else if (i == 2) {
    print(paste("getting data for ", sectors[i,2]))
    dt2 <- qcewGetIndustryData("2015", "3", sectors[i,1])    
    dt2 <- aggregate(dt2$month2_emplvl, by=list(Category=dt2$area_fips), FUN=sum)
    names(dt2) <- c("area_fips", as.character(sectors[i,2]))
    data <- merge(dt1, dt2, all=TRUE)
    
  } else {
    print(paste("getting data for ", sectors[i,2]))
    dt3 <- qcewGetIndustryData("2015", "3", sectors[i,1]) 
    dt3 <- aggregate(dt3$month2_emplvl, by=list(Category=dt3$area_fips), FUN=sum)
    names(dt3) <- c("area_fips", as.character(sectors[i,2]))
    data <- merge(data, dt3, all=TRUE)
    head(data)
      }
  
}
```

#### Assign row names and column names for reference through the transformations

```{r rename, include = FALSE}
rownames(data) <- data$area_fips
colnames(data) <- c("FIPS", "AgriForest","Mining", "Utilities","Construction",  "Manufacturing", 
                    "WholeTrade", "RetailTrade","Transportation",  "Information", "Finance", 
                    "RealEstate", "ProfScienceTech", "Management", "AdminWaste", 
                    "Education","HealthCare", "AccomodationFood","Other","PubAdmin")
glimpse(data)
str(data)
```


```{r calculateLQ, include = FALSE}
#this can only be computed for complete cases
data <- data[complete.cases(data),]

## cut out the US000 row, which is total US per sector
data <- data[data$FIPS != "US000",]

### Convert the data frame to a matrix and take a look
mat <- data.matrix( data[ ,2:length(data[1,]) ] )

# Calculate the LQ
share_tech_city <- mat / rowSums (mat)
share_tech_total <- colSums (mat) / sum (mat)
share_tech_total

lq <- t(share_tech_city) / share_tech_total
lq[is.na(lq)] <- 0 
lq[which(lq==Inf)] <- 0 
lq <- t(lq)


# here we go, this is the location quotient for each spatial unit - economic unit pair
colnames(lq) <- list("AgriForest", "Mining", "Utilities","Construction", "Manufacturing",  
                     "WholeTrade", "RetailTrade", "Transportation",  "Information", "Finance",
                     "RealEstate", "ProfScienceTech", "Management", "AdminWaste",
                     "Education","HealthCare", "AccomodationFood","Other","PubAdmin")
head(lq)
```


Data were downloaded from the Bureau of Labor Statistics, and they contain the employment levels for each county at the 2-digit NAICS level. This same analysis could be replicated at various other levels of analysis by adapting the R script that lists the NAICS codes. More information on the BLS API and data availability can be found at the [BLS website](enter here!!! )


## Factor Analysis with Verimax Rotation of LQ data
Factor analysis provides insight to the underlying relationship between multiple variables in a data set of numeric data. For this post, we analyzed the location quotients of each county in the US to see what patterns would emerge in the co-occurence of sectors. In the output below, each factor represents a composite variable based on the loadings shown in the column. Using these loadings, the factors were applied to each county using matrix multiplication to determine a score for each factor shown. The resulting score was then attributed to that county as the factor score for that variable. This method essentially simplifies the economy into 5 composite sectors. Omitted columns represent insignificant loadings, and these were also removed from the matrix multiplication. 
```{r factanal, echo=FALSE, warning=FALSE}
##########################################################
##  Perform Factor Analysis using Maximum Likelihood
##  with Varimax Rotation
##########################################################

# These are already complete cases
#lq <- lq[complete.cases(lq),]

fact1=factanal(lq[,c(1:17,19)],factors=5,rotation="varimax")
fact1

#get loading plot for factors 1 by 2-5
par(mfrow=c(3,3),mar = c(4,4,1,1))

plot(fact1$loadings, pch=18, col='red')
abline(h=0)
abline(v=0)
text(fact1$loadings, labels=colnames(lq),xpd=TRUE)

#get loading plot for factors 1 through 5
plot(fact1$loadings[,c(1,3)], pch=18, col='red')
abline(h=0)
abline(v=0)
text(fact1$loadings[,c(1,3)], labels=colnames(lq),xpd=TRUE)

plot(fact1$loadings[,c(1,4)], pch=18, col='red')
abline(h=0)
abline(v=0)
text(fact1$loadings[,c(1,4)], labels=colnames(lq),xpd=TRUE)

plot(fact1$loadings[,c(1,5)], pch=18, col='red')
abline(h=0)
abline(v=0)
text(fact1$loadings[,c(1,5)], labels=colnames(lq),xpd=TRUE)
# 4

plot(fact1$loadings[,c(2,3)], pch=18, col='red')
abline(h=0)
abline(v=0)
text(fact1$loadings[,c(2,3)], labels=colnames(lq),xpd=TRUE)

plot(fact1$loadings[,c(2,4)], pch=18, col='red')
abline(h=0)
abline(v=0)
text(fact1$loadings[,c(2,4)], labels=colnames(lq),xpd=TRUE)

plot(fact1$loadings[,c(2,5)], pch=18, col='red')
abline(h=0)
abline(v=0)
text(fact1$loadings[,c(2,5)], labels=colnames(lq),xpd=TRUE)

plot(fact1$loadings[,c(3,4)], pch=18, col='red')
abline(h=0)
abline(v=0)
text(fact1$loadings[,c(3,4)], labels=colnames(lq),xpd=TRUE)

plot(fact1$loadings[,c(3,5)], pch=18, col='red')
abline(h=0)
abline(v=0)
text(fact1$loadings[,c(3,5)], labels=colnames(lq),xpd=TRUE)
#8
#######################
```

```{r include=FALSE}
#get reproduced correlation matrix
repro1=fact1$loadings%*%t(fact1$loadings)
#residual correlation matrix -- should be close to zero
resid1=fact1$cor-repro1
round(resid1,2)

#get root-mean squared residuals
len=length(resid1[upper.tri(resid1)])
RMSR1=sqrt(sum(resid1[upper.tri(resid1)]^2)/len)
RMSR1

#get proportion of residuals greater than 0.05 in absolute value
sum(rep(1,len)[abs(resid1[upper.tri(resid1)])>0.05])/len
# smaller is better -- 0.2549 is ok but not great
```

## Factor Characterizations
To make these factors understandable, we will go through each factor, look at the loadings and name them based on what they represent.
####Factor 1: 
The first factor has relatively high scores in Produducing and Natural Resources, and relatively low scores in Services. So counties with high scores in this factor would be 

```{r createFactCols }
fal <- matrix(fact1$loadings[c(1:18),],nrow=18, ncol=5, dimnames = list(c("AgriForest",
"Mining", "Utilities",
"Construction", 
    "Manufacturing",  "WholeTrade", "RetailTrade",
"Transportation",  "Information", "Finance",
"RealEstate",  "ProfScienceTech", 
    "Management",  "AdminWaste","Education","HealthCare","AccomodationFood","PubAdmin"), 
    c("Factor1", "Factor2",   "Factor3",   "Factor4",    "Factor5")))

lqdf <- as.data.frame(lq)
lqdf$FIPS <- rownames(lqdf)

### Loop through the lq matrix and multiply 
for (i in 1:nrow(lq)) {
   lqdf$fact1[i] <- sum(data.frame(
     mapply(`*`,lqdf[ i, c(5,6,8:13,15,17,18)], fal[c(5,6,8:13,15,17,18),1],SIMPLIFY=FALSE)) )
   lqdf$fact2[i] <- sum(data.frame(
     mapply(`*`,lqdf[ i, c(5:7,10,16,17)], fal[c(5:7,10,16,17),2],SIMPLIFY=FALSE)) )
   lqdf$fact3[i] <- sum(data.frame(
     mapply(`*`,lqdf[ i, c(4,5,11,12,14,15,17)], fal[c(4,5,11,12,14,15,17),3],SIMPLIFY=FALSE)) )
   lqdf$fact4[i] <- sum(data.frame(
     mapply(`*`,lqdf[ i, c(1:8,12,14,15,17)], fal[c(1:8,12,14,15,17),4],SIMPLIFY=FALSE)) )
   lqdf$fact5[i] <- sum(data.frame(
     mapply(`*`,lqdf[ i, c(3,5,13,15,18)], fal[c(3,5,13,15,18),5],SIMPLIFY=FALSE)) )
}
```


```{r mergeCounties}
counties@data <- join(x = counties@data, y = lqdf, by = "FIPS")
```

[leaflet link](https://rstudio.github.io/leaflet/showhide.html)
```{r leaflet, warning=FALSE }
library(rgdal)
library(leaflet)
library(shiny)

pal <- colorQuantile("YlGn", NULL, n = 5)
pal2 <- colorQuantile("YlGnBu", NULL, n = 5)
pal3 <- colorQuantile("OrRd", NULL, n = 5)
pal4 <- colorQuantile("BuGn", NULL, n = 5)
pal5 <- colorQuantile("PuBu", NULL, n = 5)

county_popup <- paste0("<strong>County: </strong>", 
                      counties$NAME, 
                      "<br><strong>Factor 1: </strong>", 
                      round(counties$fact1),2)
county_popup2 <- paste0("<strong>County: </strong>", 
                      counties$NAME, 
                      "<br><strong>Factor 2: </strong>", 
                      round(counties$fact2),2)
county_popup3 <- paste0("<strong>County: </strong>", 
                      counties$NAME, 
                      "<br><strong>Factor 3: </strong>", 
                      round(counties$fact3),2)
county_popup4 <- paste0("<strong>County: </strong>", 
                      counties$NAME, 
                      "<br><strong>Factor 4: </strong>", 
                      round(counties$fact4),2)
county_popup5 <- paste0("<strong>County: </strong>", 
                      counties$NAME, 
                      "<br><strong>Factor 5: </strong>", 
                      round(counties$fact5),2)

leaflet() %>%
  setView(lng = -90, lat = 37, zoom = 3) %>% # center view on the US
  addProviderTiles("CartoDB.Positron") %>% # this sets the base layer
  addPolygons(data = counties,  # spatialPolygonDataFrame - could be lines or points
              group = "Factor 1", # this matches "base groups" below
              fillColor = ~ pal(fact1), # define the color column
              stroke = TRUE, # outline
              fillOpacity = 0.6, 
              color = "#BDBDC3", 
              weight = .5, popup = county_popup) %>%
    addPolygons(data = counties, group = "Factor 2", fillColor = ~ pal2(fact2),
                stroke = TRUE, fillOpacity = 0.6, color = "#BDBDC3", 
                weight = .5, popup = county_popup2) %>%
    addPolygons(data = counties, group = "Factor 3", fillColor = ~ pal3(fact3),
                stroke = TRUE, fillOpacity = 0.6, color = "#BDBDC3", 
                weight = .5, popup = county_popup3) %>%
    addPolygons(data = counties, group = "Factor 4", fillColor = ~ pal4(fact4),
                stroke = TRUE, fillOpacity = 0.6, color = "#BDBDC3", 
                weight = .5, popup = county_popup4) %>%
    addPolygons(data = counties, group = "Factor 5",fillColor = ~ pal5(fact5),
                stroke = TRUE, fillOpacity = 0.6, color = "#BDBDC3", 
                weight = .5, popup = county_popup5) %>%
  addLayersControl(
    baseGroups = c("Factor 1", "Factor 2", "Factor 3", "Factor 4","Factor 5"),
    #overlayGroups = c("Quakes", "Outline"),
    options = layersControlOptions(collapsed = FALSE)
  )

```