---
title: "Notes 10-26-2015"
author: "Ryan Thomas"
date: "October 26, 2015"
output: html_document
---

Testing data using lm and summary
First load the data

```{r}
Age <- c(18, 23, 25, 35, 65, 54, 34, 56, 72, 19, 23, 42, 18, 39, 37)
Max.Rate <- c(202, 186, 187, 180, 156, 169, 174, 172, 153, 199, 193, 174, 198, 
    183, 178)

plot(Max.Rate ~ Age)
```

Now summarize the relationship:
```{r}
m <- lm(Max.Rate ~ Age)
summary(m)
```

```{r}
plot(Max.Rate~Age)
abline(m)
```

Of course, ```m``` will have a lot more attributes than the ones shown in the summary.
```{r}
str(m)
```


```{r}
coef(m)
coef(m)[1]
summary(resid(m))
```

Top-left - looking for a non-relationship to confirm there is no bias in the model. 

Top-right Should be a straight line

Bottom-left Looking for outliers in the data in the form of lines that do not fit the curve.

Bottom-right Checking to see each point's influence on the model
```{r}
par(mfrow = c(2,2))
plot(m)
```


Ok, the model looks good, so let's use it to predict the 
```{r}
predict(object = m, newdata = data.frame(Age = c(50, 60)) )

plot(Age, Max.Rate)
abline(m)
ci.lwr <- predict(m, data.frame(Age = sort(Age)), level = 0.95, interval = "confidence")[, 
    2]
points(sort(Age), ci.lwr, type = "l", lty = 3)

ci.upr <- predict(m, data.frame(Age = sort(Age)), level = 0.95, interval = "confidence")[, 
    3]
points(sort(Age), ci.upr, type = "l", lty = 3)
points((predict(object = m, newdata = data.frame(Age = c(50, 60)) )), pch = 16, col = "red3")

```


